### 2023-02-13 chapters/fairness/fvgm/fairness_influence_functions.tex
fvgm appendix

For both the datasets, we consider biological `sex' as the sensitive features.
In both cases, we denote the sensitive groups `male' and `female' as `sex=0' and `sex=1', respectively. 
FIFs computed for these sensitive groups show influence of different features and their relevant disparities. We illustrate the results in Figure~\ref{fig:fif}.

Figure~\ref{fig:fif_a} illustrates that the PPV value for the `male' population in the dataset is $0.61$. If we now replace the feature `age' with uniformly random values, the PPV for the same group becomes $0.46$, Thus, the FIF of the feature `age' for `male' group is $0.15$.
Similarly, all the red bars in Figures~\ref{fig:fif_a}-\ref{fig:fif_b}, ~\ref{fig:fif_c}-\ref{fig:fif_d} indicate a positive influence of that feature and the green bars indicate a negative influence of the feature on the individual getting classified to $\hat{Y}=1$.

In Figures~\ref{fig:fif_c} and~\ref{fig:fif_f}, we illustrate the influence of different features on the disparate impact between the two groups `male' and `female'. 
The green bar indicates that removing a feature causes an increase in DI, i.e. fairness in classification, and the red bar indicates the opposite. Alternatively, we can conclude that bigger is the green bar for a feature, higher is the bias-inducing effect of it.

Thus, from Figure~\ref{fig:fif_c}, we conclude that `age' is the most bias-inducing feature among the two groups `male' and `female'.
From Fig~\ref{fig:fif_a} and~\ref{fig:fif_b}, we observe that `age' is a decisive feature for `male' while it is comparatively insignificant for `female'.

In case of Adult dataset, the PPV for `male' and `female' groups are 0.45 and 0.3 respectively. Among all the features, we observe that `race' has a positive impact on the classification for both the groups. This means removing the `race' information decreases the chance of getting classified to higher economic group. On the other hand, `capital-gain' has the highest influence on the classification for both groups. It is also the most biased-inducing feature as it varies most disparately among the two groups as the `capital-gain' differs significantly between two groups. 
\begin{figure}[t!]
	\begin{center}

		\subfloat[COMPAS]{\includegraphics[scale=0.24]{figures/dependency_exp_Learn-dependency_compas_lr_sex_0}\label{fig:fif_a}}
		\subfloat[COMPAS]{\includegraphics[scale=0.24]{figures/dependency_exp_Learn-dependency_compas_lr_sex_1}\label{fig:fif_b}}\\
		\subfloat[Adult]{\includegraphics[scale=0.24]{figures/dependency_exp_Learn-dependency_adult_lr_sex_0}\label{fig:fif_d}}
		\subfloat[Adult]{\includegraphics[scale=0.24]{figures/dependency_exp_Learn-dependency_adult_lr_sex_1}\label{fig:fif_e}}\\
		\subfloat[COMPAS]{\includegraphics[scale=0.24]{figures/dependency_exp_Learn-dependency_compas_lr_sex_DI}\label{fig:fif_c}}
		\subfloat[Adult]{\includegraphics[scale=0.24]{figures/dependency_exp_Learn-dependency_adult_lr_sex_DI}\label{fig:fif_f}}\\

	\end{center}
	\caption{Extended results on computing feature influence functions (FIF) for COMPAS  and Adult dataset.}\label{fig:fif}
\end{figure}

>>>>

For both the datasets, we consider biological `sex' as the sensitive features.
In both cases, we denote the sensitive groups `male' and `female' as `sex=0' and `sex=1', respectively. 
FIFs computed for these sensitive groups show influence of different features and their relevant disparities. We illustrate the results in Figure~\ref{fairness_fvgm_fig:fif}.

Figure~\ref{fairness_fvgm_fig:fif_a} illustrates that the PPV value for the `male' population in the dataset is $0.61$. If we now replace the feature `age' with uniformly random values, the PPV for the same group becomes $0.46$, Thus, the FIF of the feature `age' for `male' group is $0.15$.
Similarly, all the red bars in Figures~\ref{fairness_fvgm_fig:fif_a}-\ref{fairness_fvgm_fig:fif_b}, ~\ref{fairness_fvgm_fig:fif_c}-\ref{fairness_fvgm_fig:fif_d} indicate a positive influence of that feature and the green bars indicate a negative influence of the feature on the individual getting classified to $\hat{Y}=1$.

In Figures~\ref{fairness_fvgm_fig:fif_c} and~\ref{fairness_fvgm_fig:fif_f}, we illustrate the influence of different features on the disparate impact between the two groups `male' and `female'. 
The green bar indicates that removing a feature causes an increase in DI, i.e. fairness in classification, and the red bar indicates the opposite. Alternatively, we can conclude that bigger is the green bar for a feature, higher is the bias-inducing effect of it.

Thus, from Figure~\ref{fairness_fvgm_fig:fif_c}, we conclude that `age' is the most bias-inducing feature among the two groups `male' and `female'.
From Fig~\ref{fairness_fvgm_fig:fif_a} and~\ref{fairness_fvgm_fig:fif_b}, we observe that `age' is a decisive feature for `male' while it is comparatively insignificant for `female'.

In case of Adult dataset, the PPV for `male' and `female' groups are 0.45 and 0.3 respectively. Among all the features, we observe that `race' has a positive impact on the classification for both the groups. This means removing the `race' information decreases the chance of getting classified to higher economic group. On the other hand, `capital-gain' has the highest influence on the classification for both groups. It is also the most biased-inducing feature as it varies most disparately among the two groups as the `capital-gain' differs significantly between two groups. 
\begin{figure}
	\begin{center}

		\subfloat[COMPAS]{\includegraphics[scale=0.35]{figures/fairness/fvgm/dependency_exp_Learn-dependency_compas_lr_sex_0}\label{fairness_fvgm_fig:fif_a}}
		\subfloat[COMPAS]{\includegraphics[scale=0.35]{figures/fairness/fvgm/dependency_exp_Learn-dependency_compas_lr_sex_1}\label{fairness_fvgm_fig:fif_b}}\\
		\subfloat[Adult]{\includegraphics[scale=0.35]{figures/fairness/fvgm/dependency_exp_Learn-dependency_adult_lr_sex_0}\label{fairness_fvgm_fig:fif_d}}
		\subfloat[Adult]{\includegraphics[scale=0.35]{figures/fairness/fvgm/dependency_exp_Learn-dependency_adult_lr_sex_1}\label{fairness_fvgm_fig:fif_e}}\\
		\subfloat[COMPAS]{\includegraphics[scale=0.35]{figures/fairness/fvgm/dependency_exp_Learn-dependency_compas_lr_sex_DI}\label{fairness_fvgm_fig:fif_c}}
		\subfloat[Adult]{\includegraphics[scale=0.35]{figures/fairness/fvgm/dependency_exp_Learn-dependency_adult_lr_sex_DI}\label{fairness_fvgm_fig:fif_f}}\\

	\end{center}
	\caption{Extended results on computing feature influence functions (FIF) for COMPAS  and Adult dataset.}\label{fairness_fvgm_fig:fif}
\end{figure}

### 2023-02-13 tables/fairness/fvgm/fairness_metrics_on_compound_sensitive_groups.tex
fvgm appendix

\begin{figure*}[!t]
	\begin{center}
		\subfloat[]{\includegraphics[scale=0.33]{figures/sensitive_attribute__Disparate_impact_Adult_LR_Learn-efficient-dependency}}
		\subfloat[]{\includegraphics[scale=0.33]{figures/sensitive_attribute__Statistical_parity_Adult_LR_Learn-efficient-dependency}}
		\subfloat[]{\includegraphics[scale=0.33]{figures/sensitive_attribute_eqo_Adult_LR_Learn-efficient-dependency}}
		\subfloat[]{\includegraphics[scale=0.33]{figures/sensitive_attribute__PCF_Adult_LR_Learn-efficient-dependency}}	
		\\

		\subfloat[]{\includegraphics[scale=0.33]{figures/sensitive_attribute__Disparate_impact_Compas_LR_Learn-efficient-dependency}}			
		\subfloat[]{\includegraphics[scale=0.33]{figures/sensitive_attribute__Statistical_parity_Compas_LR_Learn-efficient-dependency}}
		\subfloat[]{\includegraphics[scale=0.33]{figures/sensitive_attribute_eqo_Compas_LR_Learn-efficient-dependency}}
		\subfloat[]{\includegraphics[scale=0.33]{figures/sensitive_attribute__PCF_Compas_LR_Learn-efficient-dependency}}

	\end{center}
		\caption{Verifying compound sensitive groups with respect to multiple fairness metrics. In each plot, the $ X $-axis shows sensitive features with the number of compound groups (within parenthesis) and $ Y $-axis shows computed group and causal fairness metrics. }
		\label{fig:multiple_metrics}

\end{figure*}

>>>>

\begin{figure}
	\begin{center}
		\subfloat{\includegraphics[scale=0.4]{figures/fairness/fvgm/sensitive_attribute__Disparate_impact_Adult_LR_Learn-efficient-dependency}}
		\subfloat{\includegraphics[scale=0.4]{figures/fairness/fvgm/sensitive_attribute__Statistical_parity_Adult_LR_Learn-efficient-dependency}}
		\vspace{-1em}

		\subfloat{\includegraphics[scale=0.4]{figures/fairness/fvgm/sensitive_attribute_eqo_Adult_LR_Learn-efficient-dependency}}
		\subfloat{\includegraphics[scale=0.4]{figures/fairness/fvgm/sensitive_attribute__PCF_Adult_LR_Learn-efficient-dependency}}	
		\vspace{-1em}

		\subfloat{\includegraphics[scale=0.4]{figures/fairness/fvgm/sensitive_attribute__Disparate_impact_Compas_LR_Learn-efficient-dependency}}			
		\subfloat{\includegraphics[scale=0.4]{figures/fairness/fvgm/sensitive_attribute__Statistical_parity_Compas_LR_Learn-efficient-dependency}}
		\vspace{-1em}

		\subfloat{\includegraphics[scale=0.4]{figures/fairness/fvgm/sensitive_attribute_eqo_Compas_LR_Learn-efficient-dependency}}
		\subfloat{\includegraphics[scale=0.4]{figures/fairness/fvgm/sensitive_attribute__PCF_Compas_LR_Learn-efficient-dependency}}

	\end{center}
		\caption{Verifying compound sensitive groups with respect to multiple fairness metrics. In each plot, the $ X $-axis shows sensitive features with the number of compound groups (within parenthesis) and $ Y $-axis shows computed group and causal fairness metrics. }
		\label{fairness_fvgm_fig:multiple_metrics}

\end{figure}

### 2023-02-13 tables/fairness/fvgm/sanity_check.tex
fvgm appendix

\begin{figure}
	\begin{center}
%		\subfloat[]{\includegraphics[scale=0.25]{figures/sanity_probability}\label{fig:synthetic_dist}}
		\subfloat[]{\includegraphics[scale=0.45]{figures/synthetic_example_BN}\label{fig:synthetic_bn}}

%		\subfloat[]{\includegraphics[scale=0.25]{figures/sanity_DI_LR}}
%		\subfloat[]{\includegraphics[scale=0.25]{figures/sanity_DI_SVM}}
%		\subfloat[]{\includegraphics[scale=0.25]{figures/sanity_DI_DT}}
%		
%		
%		\subfloat[]{\includegraphics[scale=0.25]{figures/sanity_SPD_LR}}
%		\subfloat[]{\includegraphics[scale=0.25]{figures/sanity_SPD_SVM}}
%		\subfloat[]{\includegraphics[scale=0.25]{figures/sanity_SPD_DT}}
%		
		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_ppv_max_PPV_LR}\label{fig:synthetic_ppv_max_LR}}
		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_ppv_max_PPV_SVM}}\\
%		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_ppv_max_PPV_DT}}
		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_ppv_min_PPV_LR}}
		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_ppv_min_PPV_SVM}\label{fig:synthetic_ppv_min_SVM}}
%		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_ppv_min_PPV_DT}\label{fig:synthetic_ppv_b}}

		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_sample_size_DI_LR}\label{fig:synthetic_sample_size_LR}}
		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_sample_size_DI_SVM}}\\
%		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_sample_size_DI_DT}}	
		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_sample_size_SPD_LR}}
		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_sample_size_SPD_SVM}\label{fig:synthetic_sample_size_SVM}}
%		\subfloat[]{\includegraphics[scale=0.33]{figures/sanity_sample_size_SPD_DT}\label{fig:synthetic_sample_size_b}}

@@

	\caption{
		Measuring accuracy of different fairness verifiers for Example~\ref{example:intro}. LR, SVM, DT refers to Logistic Regression classifier, Support Vector Machine, and Decision Tree, respectively.}
	\label{fig:synthetic_results}
\end{figure}

>>>>

\begin{figure}
	\begin{center}
		\subfloat[]{\includegraphics[scale=0.5]{figures/fairness/fvgm/synthetic_example_BN}\label{fairness_fvgm_fig:synthetic_bn}}

		\subfloat[]{\includegraphics[scale=0.4]{figures/fairness/fvgm/sanity_ppv_max_PPV_LR}\label{fairness_fvgm_fig:synthetic_ppv_max_LR}}
		\subfloat[]{\includegraphics[scale=0.4]{figures/fairness/fvgm/sanity_ppv_min_PPV_LR}\label{fairness_fvgm_fig:synthetic_ppv_min_LR}}

		\subfloat[]{\includegraphics[scale=0.4]{figures/fairness/fvgm/sanity_sample_size_DI_LR}\label{fairness_fvgm_fig:synthetic_sample_size_LR_DI}}
		\subfloat[]{\includegraphics[scale=0.4]{figures/fairness/fvgm/sanity_sample_size_SPD_LR}\label{fairness_fvgm_fig:synthetic_sample_size_LR_SPD}}
	\end{center}

	\caption{
		Measuring accuracy of different fairness verifiers for Example~\ref{fvgm_example:intro} on logistic regression classifier.}
	\label{fairness_fvgm_fig:synthetic_results}
\end{figure}

\begin{figure}
	\begin{center}
		\subfloat{\includegraphics[scale=0.4]{figures/fairness/fvgm/sanity_ppv_max_PPV_SVM}}
		\subfloat{\includegraphics[scale=0.4]{figures/fairness/fvgm/sanity_ppv_min_PPV_SVM}}\\
		\subfloat{\includegraphics[scale=0.4]{figures/fairness/fvgm/sanity_sample_size_DI_SVM}}
		\subfloat{\includegraphics[scale=0.4]{figures/fairness/fvgm/sanity_sample_size_SPD_SVM}}

@@

	\caption{
		Measuring accuracy of different fairness verifiers for Example~\ref{fvgm_example:intro} on SVM classifier.}
	\label{fairness_fvgm_fig:synthetic_results_SVM}
\end{figure}
